---
title: "Final Project"
author: "Albina Music"
date: "2023-05-15"
categories: [Final]
execute: 
  error: true
---

## Title:

# Research Question - Do you come across as smarter when people read what you say or hear what you say?

## Lab skills learned
1. Conduct independent samples t-tests
2. Generate figures
3. Discuss the results and implications

# Load Libraries
```{r}
library(data.table)
#all_data <- fread("https://raw.githubusercontent.com/CrumpLab/statisticsLab/master/data/SchroederEpley2015data.csv")

all_data <- fread("article.csv") 
```


```{r}
library(tidyverse)
library(summarytools)
```

## Obtain openly available data from an existing primary source article. For example, psychology papers may include links to open data from the published experiment.

## Here are a few tips for finding a psych paper with open data. Most important, for this assignment you do not need to re-analyze all of the data from a particular paper. Many papers have multiple experiments, and multiple analyses, including analyses you may not be familiar with. You can restrict your re-analysis to a portion of the paper that involves a data-visualization of interest.

https://osf.io The open science framework contains many repositories of open data that are part of published papers

## Some journals, including Psychological Science, put badges on papers with open data. Look for the blue open-data badge. You will usually find the link to the open-data in the paper.

## Import, pre-process, and visualize the data using R. After you have obtained the data, your main task is to produce a visualization of the data. For example if the authors provided a graph of their data, can you recreate this graph in your own analysis.

## Write a brief report about your visualization. Include a brief description of the research question and experiment (with citation to the paper, and link to find the data). Include the R code chunks necessary to complete the data-visualization. Include a brief description or explanation of the pattern of results in the visualization, and how they relate to the research question.

## Publish your completed report on your blog.


